---
title: " GE=Meat to Chastity Adjusted (Irina, OPSB-based) DEGs SEGs from NPSB; (ii) make standard format for both OPSB and NPSB DEG and SEG sets, both R & Matlab (no 'set' column for Matlab)"
output:
  word_document: default
  html_notebook: default
---

-libraries
```{r}
#sh be installed tidyverse,reshape2,ggplot2,ggpubr
library(reshape2)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(tidyr)
library(ggpubr)
#library("readxl")
library(datasets)  
library(purrr)
```

- get new PSB raw data, filtered and modified
```{r}

param3='Christel NPSB'

source("new_PSB.R")
main_raw <-new_PSB()
dim (main_raw)->size_mr
size_mr

#read_tsv("C:/Users/ia1/Documents/R_work/Ect_Mes_DEG/E7_5_pseudobulk_log2CPM_200331.txt", col_types = cols(Chromosome = col_character())) -> expressionRaw
#expressionRaw
#main_raw

```

- read Chasity adjusted (based on OPSB, Irina) unique set


```{r}
param5="Chastity adj unique"
source("readCh_adj_Dif_Sim.R")
readCh_adj_Dif_Sim()->gcha

```


#- read Dif Christels unique sets, high

```{r}
#-------------------New Diff Genes: separate Ect 
#param5='Ch new unique DEGs, high'

#print(param5)

```
- define in one go from read files

```{r}
dect=gcha[[1]]
#dect
dend=gcha[[2]]
#dend
dmes=gcha[[3]]
#dmes
sim=gcha[[4]]
#sim
dect

```
- visualise DEG SEG old PSB

```{r}
source("densityGE_any.R")

DS_oldPSB<-rbind(dect,dend,dmes,sim)

peems <-densityGE_any(DS_oldPSB)
peems + facet_wrap( ~ set)+ ggtitle(" with old PSB GE")


  
```
- box plot per DEG set
```{r}
boxplot(dect[6:8])
```

- paired t-tes
```{r}
#------------run paired t0test

t.test(dect$ecto, dect$meso, paired=T)


```

- box plots of sets
```{r}
#boxplot(working.memory[2:3])
```

- save/write R OSPB format
```{r}


# dmes %>%
#   write_tsv("DifMes_OPSB_R_Cha.txt")
# 
# dend %>%
#   write_tsv("DifEnd_OPSB_R_Cha.txt")
# 
# dect %>%
#   write_tsv("DifEct_OPSB_R_Cha.txt")
# 
# sim %>%
#   write_tsv("Sim_OPSB_R_Cha.txt")


```
- save/write Matlab format OPSB
```{r}


#dmes %>%
 # select(-set,-strand)%>%
 # write_tsv("DifMes_OPSB_mtlb_Cha.txt")

#dend %>%
 # select(-set,-strand)%>%
 # write_tsv("DifEnd_OPSB_mtlb_Cha.txt")

#dect %>%
 # select(-set,-strand)%>%
 # write_tsv("DifEct_OPSB_mtlb_Cha.txt")

#sim %>%
 # select(-set,-strand)%>%
 # write_tsv("Sim_OPSB_mtlb_Cha.txt")

```

- extract DEG vectors from DEGs SEGs to grab  new GE later
 for NPSB
  
```{r}
#----------------submit  DEGs, extract dEG vectors


dect %>% 
  distinct(gene) %>% 
  as_vector() -> DEct_vector
#DEct_vector
len_Dec <- length(DEct_vector)
len_Dec

dend %>% 
  distinct(gene) %>% 
  as_vector() -> DEnd_vector
#DEnd_vector
len_Den <- length(DEnd_vector)
len_Den


dmes %>% 
  distinct(gene) %>% 
  as_vector() -> DMes_vector

len_Dm <- length(DMes_vector)
len_Dm

#------------------Sim
sim %>% 
  distinct(gene) %>% 
  as_vector() -> Sim_vector

len_Ds <- length(Sim_vector)
len_Ds


```
- collect expression for DEG sets from main_raw data NPSB

```{r}

main_raw %>%
  rename(strand=Strand)%>%
    mutate(set=case_when(gene %in% DEct_vector ~"DifEc",
                         gene %in% DEnd_vector ~"DifEn",
                         gene %in% DMes_vector ~"DifM",
                         gene %in% Sim_vector ~"Sim")
    )->separated_DS

#separated_DS


```

- split due to sets with new PSB GE
```{r}

  sett<-separated_DS %>%
group_by(set)

group_split(sett)->ss
#ss

#--------------------------------

DifEct_NPSB <-ss[[1]]
DifEct_NPSB %>%
    rename(chr=Chromosome,start=Start,end=End,ecto=Ecto, endo=Endo,meso=Meso)->DifEct_NPSB

DifEnd_NPSB <-ss[[2]]
DifEnd_NPSB %>%
    rename(chr=Chromosome,start=Start,end=End,ecto=Ecto, endo=Endo,meso=Meso)->DifEnd_NPSB


DifMes_NPSB <-ss[[3]]

DifMes_NPSB%>%
  rename(chr=Chromosome,start=Start,end=End,ecto=Ecto, endo=Endo,meso=Meso)->DifMes_NPSB


Sim_NPSB <- ss[[4]]
Sim_NPSB%>%
    rename(chr=Chromosome,start=Start,end=End,ecto=Ecto, endo=Endo,meso=Meso)->Sim_NPSB
#Sim_NPSB
Comm_notCha <-ss[[5]]
Comm_notCha%>%
    rename(chr=Chromosome,start=Start,end=End,ecto=Ecto, endo=Endo,meso=Meso)->Comm_notCha_NPSB

```
- Mtalab format OPSB
```{r}
DifEct_NPSB%>%
  select(gene,chr,start,end,ecto,endo,meso)->DifEct_NPSB_mtlb_Cha
DifEnd_NPSB%>%
  select(gene,chr,start,end,ecto,endo,meso)->DifEnd_NPSB_mtlb_Cha
DifMes_NPSB%>%
  select(gene,chr,start,end,ecto,endo,meso)->DifMes_NPSB_mtlb_Cha
 Sim_NPSB%>%
  select(gene,chr,start,end,ecto,endo,meso)->Sim_NPSB_mtlb_Cha
     

```

 - visualise DEA SEG new PSB

```{r}
source("densityGE_any.R")

EcEnMS<-rbind(DifEct_NPSB,DifEnd_NPSB,DifMes_NPSB,Sim_NPSB)

pecenms<-densityGE_any(EcEnMS)
pecenms + facet_wrap( ~ set) + ggtitle(" with new, NPSB GE")


```
- save/write R NSPB format
```{r}


#DifMes_NPSB %>%
  #write_tsv("DifMes_NPSB_R_Cha.txt")

#DifEnd_NPSB %>%
  #write_tsv("DifEnd_NPSB_R_Cha.txt")

#DifEct_NPSB %>%
  #write_tsv("DifEct_NPSB_R_Cha.txt")

#Sim_NPSB %>%
  #write_tsv("Sim_NPSB_R_Cha.txt")

```
- save/write mtlb format
```{r}


#DifMes_NPSB_mtlb_Cha %>%
 # write_tsv("DifMes_NPSB_mtlb_Cha.txt")

#DifEnd_NPSB_mtlb_Cha %>%
 # write_tsv("DifEnd_NPSB_mtlb_Cha.txt")

#DifEct_NPSB_mtlb_Cha %>%
 # write_tsv("DifEct_NPSB_mtlb_Cha.txt")

#Sim_NPSB_mtlb_Cha %>%
  #write_tsv("Sim_NPSB_mtlb_Cha.txt")

```


